### 2018-12-07 v1-task-relevance-cspc.tex
First Draft

\documentclass[english,,man,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Contextual recruitment of selective attention can be updated via changes in task-relevance},
            pdfauthor={Nicholaus P. Brosowsky~\& Matthew J. C. Crump},
            pdfkeywords={Attention, Memory, Cognitive Control, Conflict Adaptation,
Context-specific, Task Relevance},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[shorthands=off,main=english]{babel}
\else
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\fi
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

  \title{Contextual recruitment of selective attention can be updated via changes
in task-relevance}
    \author{Nicholaus P. Brosowsky\textsuperscript{1}~\& Matthew J. C.
Crump\textsuperscript{2}}
    \date{}

\shorttitle{Contextual recruitment of selective attention}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} The Graduate Center of the City University of New York\\\textsuperscript{2} Brooklyn College of the City University of New York}
\keywords{Attention, Memory, Cognitive Control, Conflict Adaptation, Context-specific, Task Relevance\newline\indent Word count: 5946}
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

\usepackage{longtable}
\usepackage{lscape}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage[flushleft]{threeparttable}
\usepackage{threeparttablex}

\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

\authornote{

Correspondence concerning this article should be addressed to Nicholaus
P. Brosowsky, The Graduate Center, CUNY, 365 5th Ave, New York, NY
10016. E-mail:
\href{mailto:nbrosowsky@gradcenter.cuny.edu}{\nolinkurl{nbrosowsky@gradcenter.cuny.edu}}}

\abstract{
Evidence across a wide variety of attention paradigms shows that
environmental cues can trigger adjustments to ongoing priorities for
attending to relevant and irrelevant information. This context-specific
control over attention suggests that cognitive control can be both
automatic and flexible. For instance, in selective attention tasks,
congruency effects are larger for items that appear in a context
associated with infrequent conflict than in a context associated with
frequent conflict. Since the to-be-presented context cannot be predicted
or prepared for in advance, attention is assumed to be rapidly updated
on-the-fly, triggered by the currently presented context.
Context-specific control exemplifies how learning and memory processes
can influence attention to enable cognitive flexibility. However, what
determines the use of previously learned associations still remains
unclear. In the current study, we examined whether task-relevance would
influence the learning and use of context cues in a flanker task. Using
a secondary counting task, context dimensions associated with differing
levels of conflict were made task-relevant or -irrelevant across the
experiment. In short, we found that making new contextual information
task-relevant caused participants to ignore a previously learned
context-attention association and adopt a new context-specific control
strategy; all without changing the experimental stimuli. This result
suggests that task-relevance is a key determinant of context-specific
control.

}

\begin{document}
\maketitle

\section{Introduction}\label{introduction}

Selective attention processes are commonly investigated using
interference paradigms such as the Stroop (1935) and flanker (B. A.
Eriksen \& Eriksen, 1974) tasks, where participants identify a target
while ignoring a response-congruent or -incongruent distractor.
Performance is typically better on congruent versus incongruent trials
and the difference-the congruency effect-taken as an index of
attentional priorities. Large congruency effects are thought to reflect
ineffective filtering of the distracting stimuli whereas small
congruency effects are thought to reflect effective filtering. By
probing factors that systematically alter congruency effects, we can
then make inferences about processes that control attentional filtering.
For example, manipulating the frequency of conflict via the proportion
of congruent versus incongruent trials has shown to influence the size
of the congruency effect. Typically, a high proportion congruent
experiment produces large congruency effects, whereas a low proportion
congruent experiment produces small congruency effects (Logan \&
Zbrodoff, 1979; Lowe \& Mitterer, 1982; West \& Baylis, 1998). This
result is usually explained as strategic control, where participants
increase attentional control under high-conflict demands and relax
attentional control under low-conflict demands (Logan, 1980; Logan \&
Zbrodoff, 1979; Logan, Zbrodoff, \& Williamson, 1984; Lowe \& Mitterer,
1982). Recent work however, has demonstrated that attentional control is
not only adjusted by top-down regulation, but can also be triggered
automatically by environmental cues (Brosowsky \& Crump, 2018; J. Bugg
\& Crump, 2012; Egner, 2014; Fischer \& Dreisbach, 2015; King, Korb, \&
Egner, 2012; Mayr \& Bryck, 2007).

For example, Crump, Gong, and Milliken (see also, Corballis \& Gratton,
2003; Crump, Gong, \& Milliken, 2006) presented Stroop stimuli in one of
two randomly chosen locations and manipulated the frequency of conflict
associated with each location. One location was associated with a high
frequency of conflict (25\% congruent trials) and the other with a low
frequency of conflict (75\% congruent trials). Overall, the proportion
of congruent trials was 50\% and randomized such that the upcoming
location could not be predicted. Even so, congruency effects were shown
to be smaller for trials where the stimulus appeared in the high
conflict location as compared to the low conflict location. This effect,
known now as the context-specific proportion congruent effect (CSPC),
has now been replicated in a number of different selective attention
paradigms (e.g., Alards-Tomalin, Brosowsky, \& Mondor, 2017; Blais,
Harris, Sinanian, \& Bunge, 2015; J. Bugg, 2014; Crump, 2016; Crump,
Milliken, Leboe-McGowan, Leboe-McGowan, \& Gao, 2018; Fischer,
Gottschalk, \& Dreisbach, 2014; Hübner \& Mishra, 2016).

Critical evidence however, that CSPC effects reflect context-specific
control rather than other non-control learning processes (e.g., Schmidt
\& Besner, 2008), comes from work showing that CSPC effects can transfer
to frequency unbiased items (Brosowsky \& Crump, 2016; Crump \&
Milliken, 2009; though, see Hutcheon \& Spieler, 2017; Weidler \& Bugg,
2016; Weidler, Dey, \& Bugg, 2018). Crump and Milliken (2009), for
example, divided Stroop items into two mutually exclusive sets (e.g.,
RED/GREEN and BLUE/YELLOW). One set was defined as the frequency biased
set, and presented with 75\% congruency in one location, and 25\%
congruency in the other. The second set however, was presented with 50\%
congruency in both locations. Nevertheless, they found smaller
congruency effects for unbiased items presented in the high conflict
location as compared to the low conflict location.

One explanation for CSPC effects is that the repeated application of
attentional priorities in a particular context creates an associative
link in episodic memory between the attentional control procedures and
the contextual information (Abrahamse, Braem, Notebaert, \& Verguts,
2016; Brosowsky \& Crump, 2018; Crump, 2016; Egner, 2014). Once the
associative link is established, processing the context is assumed to
trigger the retrieval of previous experiences, automatically reinstating
the associated attentional priorities. Although this explanation seems
plausible, it remains largely unclear how such associations are
established and perhaps more importantly, it remains unclear what
determines which learned associations will be used to guide attention at
any given moment.

The importance of the latter issue becomes increasingly clear if we
speculate about context-specificity in more complex, real-world
environments. Consider myself, right now, writing this article in a
coffee shop. The coffee shop is a complex environment that contains a
multitude of contextual cues (e.g., pictures on the wall, chairs,
coffee, laptop), each of which I have had many different experiences
applying attentional control. Although my laptop might cue attentional
processing helpful for finishing this article, other features of the
coffee shop might cue other, unhelpful attentional processing related to
other experiences like socializing and people-watching. This is all to
say that our environments, and experiences within them, afford many
different associative relationships between cognitive processing and
environmental cues. What then determines which associative relationships
will be used to guide attention (or even if such associative
relationships are used at all)?

Task-relevance might be one factor that is important for establishing
and using context-attention associations. This proposal might seem
counter-intuitive as it is typically assumed that the context, within
selective attention tasks at least, is completely irrelevant to the
task. In fact, referring to such stimulus dimensions as the
\enquote{context} carries with it an implication that the feature is
background information, irrelevant to the task-at-hand, and should be
ignored by the participant. A closer look at CSPC designs however,
suggests that the context cues may not be as irrelevant as implied. For
example, the most commonly used context cue is location (Brosowsky \&
Crump, 2016; Corballis \& Gratton, 2003; Crump, 2016; Crump et al.,
2006; Crump, Vaquero, \& Milliken, 2008; Weidler \& Bugg, 2016; Weidler
et al., 2018). Although in some sense, the location is irrelevant for
identifying the color of a word, it seems disingenuous to consider it
completely task-irrelevant. In order to identify the target, the
participant must first identify and orient to its location on the
screen. In this is sense, it is relevant to their ability to complete
the task and furthermore, location information may receive priority
during encoding (Logan, 1998; Mayr, 1996). Similar arguments could be
made about other context features that are inherent to processing the
target stimulus, like font-type (J. Bugg, Jacoby, \& Toth, 2008) and
color (Vietze \& Wendt, 2009).

To date, we are aware of only one study that has looked at
task-relevance within the CSPC literature. Crump, Vaquero, \& Milliken
(2008) used a Stroop prime-probe task where the distractor word is
presented first followed by a color patch probe. Critically, they used
the shape of the color patch probe, either a square or a circle, as the
contextual dimension; one associated with a high frequency of conflict,
the other with a low frequency. In one experiment participants were made
aware of the shape-proportion contingencies but given no additional
instructions. In the second, participants were given a secondary task to
count the number of trials that contained a square. The first experiment
failed to find evidence for context-specificity, whereas the second did.
Although not testing task-relevance directly, (2013)) found similar
effects. In this study they used images of male and female faces as
context cues in a flanker task. When given instructions to think of the
faces as members of the gender categories they found CSPC effects, but
when given instructions to think of the faces as individuals they did
not. These results suggest that making a context dimension relevant to
the on-going task is important for producing CSPC effects.

\section{The Current Study}\label{the-current-study}

Some prior suggests that the task-relevance of contextual information is
important for establishing new associations between context cues and
attentional priorities. Simply put, CSPC effects emerge when context
cues are somehow made relevant to the task, but do not when made
irrelevant (Cañadas et al., 2013; Crump et al., 2006, 2008). However, it
still remains unclear how task-relevance may or may not inform the use
of previously learned associations. In the current study, we examined
whether manipulating task-relevance would cause participants to ignore a
previously learned association and adopt a new context-specific control
strategy.

A flanker task was used to measure attentional control in an adapted
CSPC design (Crump \& Milliken, 2009; Crump, Brosowsky, \& Milliken,
2017). Similar to previous studies, flanker stimuli were presented in
contexts associated with differing levels of conflict (0\%, 50\%, or
100\% congruent). However, unlike previous CSPC designs where contexts
were defined by a single discriminating feature (e.g., upper vs.~lower
screen locations), contexts were defined by two feature dimensions (see
Figure 1A). In one condition, for example, contexts were defined by
object identity (hat or chair) and color (blue or green). Only three out
of the four possible feature combinations were presented, each
associated with a different proportion of congruent trials (0\%, 50\%,
or 100\% congruent). Critically, the frequency unbiased context (50\%
congruent) always shared a feature with each of the frequency-biased
contexts (0\% and 100\% congruent). The frequency biased contexts (0\%
and 100\% congruent) however, did not share any features (see Figure 1C
for an example).

A secondary counting task was used to manipulate the task-relevance of
the context dimensions (Crump \& Milliken, 2009) which critically,
switched halfway through the experiment. Participants were instructed to
keep a running count of one of the overlapping features (e.g.,
\enquote{count whenever a hat is presented}) which was associated with
either a high or low frequency of conflict. Halfway through the
experiment they received new instructions to count the other overlapping
feature (e.g., \enquote{count whenever a green item is presented}).
Critically, the set of context images remained the same throughout the
experiment (see Figure 1C for an example) and the task-relevance of the
context dimensions was the only aspect of the experiment that changed
from the first to second phases.

The critical measure of interest is the congruency effect produced in
the frequency unbiased context. If, on the one hand, task-relevance has
no impact on the use of context-attention associations we would expect
no differences between the congruency effects when the high-conflict
context dimension is made task-relevant or the low-conflict context
dimension is made task-relevant. This could occur because all three
contexts are treated as individual contexts throughout the whole
experiment (e.g., Cañadas et al., 2013). Or it could occur because
associations formed in the first block interfere with learning new
associations in the second (e.g., Brosowsky \& Crump, 2016). On other
hand, task-relevance might not only dictate how associations are formed,
but also whether learned associations are used. In this case, we would
expect that there will be smaller congruency effects when the
high-conflict context dimension is made task-relevant as compared to
when the low-conflict context dimension is made task-relevant. This
would demonstrate that participants were able to ignore a previously
learned association and adopt a new context-specific control strategy.

A secondary goal was to conceptually replicate previous findings and
test the generalizability of any task-relevance effects. Therefore, we
included three conditions that differed only in the kinds of stimuli
used to create context cues. In one condition we used face images where
context dimensions were defined by social categories, similar to Cañadas
et al., (2013). However, they found that social categorization occurred
spontaneously and it was unclear what impact that would have on the
current results. Therefore, we included another condition that used
images of simple objects where context dimensions were defined by object
features. Finally, Cañadas et al.(2013) also found that context effects
generalized to novel face images. In our third condition we used
non-repeating face exemplars to examine whether any task-relevance
effects would also generalize in our design or if learning would be
image-specific.

\section{Methods}\label{methods}

\begin{figure}
\centering
\includegraphics{images/Task.jpg}
\caption{\label{fig:figure1}Illustration of the stimuli and trial construction. Figure
1A shows the feature dimensions for each of the conditions. Figure 1B
shows an example trial stimulus containing both the context and flanker
images. Figure 1C shows an example of how feature dimensions could have
been assigned to each proportion congruency (PC) condition, and an
example of the secondary task assignments. The secondary task order, as
well as the feature and congruency assignments were all randomized for
each participant.}
\end{figure}

\subsection{Participants}\label{participants}

All participants were recruited from Amazon Mechanical Turk (AMT) and
compensated \$2.00 for participating. The amount compensated was
calculated by estimating the maximum amount of time required to complete
each experiment and multiplying by \$6.00 per hour. For each experiment
the number of HITs (Human intelligence tasks, an Amazon term for a
work-unit) refers to the number of participants who initiated the study.
Participants were included in the study if they completed all trials. We
posted 150 HITs and 145 participants completed all trials.

\subsection{Apparatus \& Stimuli}\label{apparatus-stimuli}

The experiments were programmed using JavaScript, CSS and HTML. The
program allowed participants to complete task only if they were running
Safari, Google Chrome, or Firefox web browsers. Flanker stimuli
consisted of images of five arrows pointed left or right presented at
250 x 50 pixels (each arrow was 50 x 50 pixels). Context stimuli were
constructed using images selected from Brady, Konkle, Gill, Oliva, and
Alvarez (2013) color-rotated to blue and green, and face images from the
Chicago Face Database (Ma, Correll, \& Wittenbrink, 2015), supplemented
with the NimStim Set of Facial Expressions (Tottenham et al., 2009). The
object images were displayed at 250 x 250 pixels, while the face images
were displayed at 250 x 313 pixels. The experiment ran as a pop-up
window that filled the entire screen. The background was white, and
stimuli were presented in the center of the screen.

Design. Experiment 1 used a 2x2x3 mixed design with task-relevant
context (0\% and 100\%) and unbiased-item congruency (congruent and
incongruent) as within-subject factors, and context-type (object,
social, and social/non-repeating) as the between-subjects factor. All
three conditions were constructed using the same general method. The
experiment was divided into two phases. Each phase consisted of 144
flanker trials (48 trials per context), and 13 count response trials for
a total of 314 trials. On the count response trials, participants
indicated how many trials they had counted until that point. The count
response trials occurred once for every 12 flanker trials and was
randomly inserted between trial 6 and 12 of each 12-trial block. Each
phase ended with one additional count response trial.

On every flanker trial, participants were presented with flanker stimuli
paired with one of three contexts. Each context was associated with a
different proportion congruency such that two cues were associated with
a biased frequency (0\% and 100\% proportion congruency), while one was
associated with an unbiased frequency (50\% proportion congruency). The
feature dimensions and corresponding context images assigned to each of
the biased and unbiased item sets were randomly determined for each
participant. However, context images used for the frequency biased
trials never shared features, while the frequency unbiased context image
always shared a feature with each of the frequency biased context images
(see Figure 1). Additionally, the feature assignments remained the same
throughout phases 1 and 2, and critically, the only change to the task
was which feature the participant was instructed to count (see Figure 1C
for an example).

All critical aspects of the task were randomized between participants.
This includes the three chosen context images, the features assigned to
proportion levels, the features assigned to each counting condition, the
secondary task order, and the order of trials.

\subsection{Procedure}\label{procedure}

All participants were AMT workers who found the experiment using the AMT
system. The participant recruitment procedure and tasks were approved by
the Brooklyn College Institutional Review Board. Each participant read a
short description of the task and gave consent by pressing a button
acknowledging they had read the displayed consent form. Participants
then completed a short demographic survey, and proceeded to the main
task, which was displayed as a pop-up window. Participants were
instructed to identify the direction of the center arrow on each trial
as quickly and accurately as possible by pressing \enquote{z} if the
arrow pointed left, and \enquote{m} if the arrow pointed right.
Additionally, they were instructed to silently keep count of the number
of trials that contained a feature. In the object context condition,
they were asked to count trials that contained a certain color (blue or
green) or object-identity (hat or chair) and in the social context
conditions they were asked to count the number of trials that contained
appeared a certain gender (male or female) or race (black or white).
Periodically throughout the experiment, participants were asked to
report how many trials they had counted until that point and to restart
their count from 0. Halfway through the experiment participants received
new instructions about which feature to count (see Figure 1C). Each
trial began with a blank inter-stimulus interval (ISI) of 400 ms,
followed by a fixation cross presented in the center of the screen for
200 ms, then a second blank ISI of 400 ms. Next, the flanker and context
stimuli appeared in the center of screen (the flanker above the context
image; see Figure 1B) and remained on screen until a response was made.
Following a response, accuracy feedback was presented for 1000 ms. A
response automatically triggered the next trial. Halfway through the
experiment (157 trials), participants received new instructions about
which feature to count and to press the button on-screen when they were
ready to continue.

\section{Results}\label{results}

\begin{figure}
\centering
\includegraphics{v1-task-relevance-cspc_files/figure-latex/figure2-1.pdf}
\caption{\label{fig:figure2}Results from experiment 1 showing congruency effects
(incongruent - congruent) for frequency unbiased items in reaction times
(left) and error rates (right) as a function of the task-relevant
context (100\% proportion congruent versus 0\% proportion congruent).}
\end{figure}

Participants with mean error rates greater than 25\% were excluded from
the analyses, eliminating 13 participants. For all remaining
participants, the correct RTs from frequency unbiased trials in each
condition were submitted to an outlier removal procedure. The
non-recursive Van Selst and Jolicoeur outlier removal procedure was
applied after removing response times greater than 3000 ms (Van Selst \&
Jolicoeur, 1994). This procedure removed 3.39\% of the total
observations.

The primary question of interest was whether the task-relevance of
context features associated with different levels of conflict would
influence the size of the congruency effect for frequency unbiased items
(see Table 1). To that end, mean correct RTs and mean error rates from
frequency unbiased trials were submitted to a mixed analysis of variance
(ANOVA) with task-relevant context (0\% and 100\% PC) and unbiased-item
congruency (congruent and incongruent) as within-subject factors, and
context-type (object, social, and social/non-repeating) as the
between-subjects factor.

The results of the RT analysis revealed the critical, two-way
interaction between task-relevant context and unbiased-item congruency
to be significant, \(F(1, 128) = 8.30\), \(\mathit{MSE} = 1,608.54\),
\(p = .005\), \(\hat{\eta}^2_p = .061\), demonstrating smaller
congruency effects when the context dimension associated with high
conflict was made task-relevant. Additionally, the three-way interaction
between the task-relevant context, unbiased-item congruency, and
context-type was non-significant, \(F(2, 128) = 0.00\),
\(\mathit{MSE} = 1,608.54\), \(p = .996\), \(\hat{\eta}^2_p = .000\),
showing no evidence for differences between the CSPC effects across
context-type manipulations.

Similarly, the corresponding error analysis also resulted in a
significant two-way interaction between the task-relevant context and
unbiased-item congruency, \(F(1, 128) = 8.30\),
\(\mathit{MSE} = 1,608.54\), \(p = .005\), \(\hat{\eta}^2_p = .061\),
and a non-significant three-way interaction between task-relevant
context, unbiased-item congruency, and context-type,
\(F(2, 128) = 0.00\), \(\mathit{MSE} = 1,608.54\), \(p = .996\),
\(\hat{\eta}^2_p = .000\). Therefore, the congruency effects in error
rates were also smaller when the context dimension associated with high
conflict was made task-relevant, corroborating the results of the
reaction time analysis.

\section{General Discussion}\label{general-discussion}

The aim of the current study was to investigate whether task-relevance
plays a role in the contextual recruitment of selective attention.
Specifically, we tested whether manipulating the relative task-relevance
of context cues could cause participants to ignore a previously learned
context-association and apply a new association. We created a frequency
unbiased context cue that shared features with two frequency biased
contexts and included a feature-counting task to manipulate the
task-relevance of context dimensions across two blocks of trials.
Critically, halfway through the experiment participants received new
instructions changing the task-relevant feature from one frequency
biased cue to the other.

The key finding was that the congruency effects for the frequency
unbiased items were significantly larger when the low-conflict context
was made task-relevant as compared to when the high-conflict context was
made task-relevant. This result is consistent with prior CSPC effects
and, like the previous work, suggests that the context cues triggered
rapid adjustments to attentional control (Crump \& Milliken, 2009).
However, unlike prior studies, we were able to experimentally manipulate
the CSPC effect across blocks of trials without changing any of the
physical properties of the stimuli. This novel finding demonstrates that
participants were able to learn and apply one context-attention
association in the first phase, and subsequently ignore that association
to learn a new association in the second phase.

This result implicates an important role for task-relevance in producing
CSPC phenomena. Crump et al. (2008) used shapes as context cues in a
prime-probe Stroop task and did not find CSPC effects until the context
cues were made task-relevant. Similarly, Cañadas et al. (2013)
eliminated the CSPC effect by making the contextual cue effectively
unrelated to the task. These studies suggest that task-relevance plays
an important role in establishing associations between contextual
information and attentional priorities to produce CSPC effects. Our
finding extends this work in two important ways. First, we show that
changing the task-relevance of the presented cues corresponded with a
change in attentional control in the predicted direction. This
demonstrates that task-relevance is also a key determinant of whether a
previously learned attention-context association will be used or ignored
in favor of a new association. Second, we show that task-relevance
allowed participants resolve competition between two competing
contextual cues, responding on the basis of one at the expense of the
other. To our knowledge, this is the first demonstration that CSPC
effects can be produced when there are multiple, overlapping contextual
cues available.

In light of prior work, we take this as evidence that the contextual
recruitment of selective attention, although likely implicit, is not
obligatory (e.g., Brosowsky \& Crump, 2016), requiring that
environmental information be incorporated into the task representation.
Similarly, there appears to be flexibility in which environmental
features are selected and used to guide attention, which can be rapidly
updated depending on the task-relevance of those features. Such a result
lends some insight into how context-specific control might operate in
more complex, real-world environments, where there is an over-abundance
of environmental features that afford many different learned
associations. From a theoretical perspective, this result is consistent
memory-based accounts of CSPC phenomena (Brosowsky \& Crump, 2018; J.
Bugg \& Hutchison, 2013; Crump \& Milliken, 2009; Crump et al., 2018,
2017). Under this view, a memory process encodes attentional priorities
in the representation of individual experiences and, as a result, become
associated with the environment where they were used. The subsequent
reoccurrence of a prior context triggers the retrieval and reinstatement
of those attentional priorities. Our results show however, that all the
features of the environment may not be treated equally and that only
task-relevant features are used to probe memory and guide attention.

Another key result of this study concerned the different stimuli used as
context cues. Across the three conditions, we varied the type of context
image and dimensions. We manipulated the type of image presented,
including both objects (identity and color dimensions) and faces (gender
and racial dimensions). We also manipulated whether a single set of
three repeating images were presented (object and social) or a set of
non-repeating images were presented (social/non-repeating). Across all
three conditions, we found no evidence that using different stimuli had
an influence on the size or direction of the CSPC demonstrating
generalizability of this phenomenon. Furthermore, CSPC effects were
present even when using non-repeating images which suggests that
context-dependency did rely on image-specific associations but
higher-order, learned categorical information.

Traditional models of person perception posit that social categories are
automatically activated in the presence of social stimuli (e.g., Brewer,
1988; Devine, 1989; Fiske \& Neuberg, 1990). Cañadas et al. (2013)
however, found that directing participants to think about faces in terms
of individual features eliminated context-specific attention effect and
suggested that momentary motivations may influence the automaticity of
social categorization. Our results add to this literature by observing
the influence of momentary motivations (i.e., task-relevance) when there
is competition between two salient social categories. Specifically, we
found that participants could categorize on the basis of one social cue
at the expense of the other, and subsequently switch between them. These
findings may speak to issues of automaticity in social categorization
(Macrae \& Bodenhausen, 2001) as well as understanding how the
situational context can prime one social identity over another (Crisp \&
Hewstone, 2007). Furthermore, we found no evidence for differences
between the social and object conditions. This suggests that
categorization within the CSPC task is quickly and easily learned when
the task supports such learning and is non-unique to social stimuli.

In sum, our results provide new evidence that changes in task-relevance
can update the contextual recruitment of selective attention in a CSPC
flanker task. Importantly, we show that making one feature task-relevant
produced a CSPC effect, even in the presence of a competing contextual
cue. Moreover, changing the task-relevant context-cue midway through the
experiment was accompanied by predictable changes in the congruency
effects. This effect was found to be generalizable across two different
kinds of stimuli and occurred even when using non-repeating images.
These results implicate an important role for task-relevance in
producing context-dependency and are consistent with memory-based
accounts of CSPC phenomena.

\newpage

\section{References}\label{references}

\begingroup
\setlength{\parindent}{-0.5in} \setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\hypertarget{ref-abrahamse_grounding_2016}{}
Abrahamse, E., Braem, S., Notebaert, W., \& Verguts, T. (2016).
Grounding cognitive control in associative learning. \emph{Psychological
Bulletin}, \emph{142}, 693--728.
doi:\href{https://doi.org/10.1037/bul0000047}{10.1037/bul0000047}

\hypertarget{ref-alards-tomalin_auditory_2017}{}
Alards-Tomalin, D., Brosowsky, N. P., \& Mondor, T. A. (2017). Auditory
statistical learning: Predictive frequency information affects the
deployment of contextually mediated attentional resources on perceptual
tasks. \emph{Journal of Cognitive Psychology}, \emph{0}(0), 1--11.
doi:\href{https://doi.org/10.1080/20445911.2017.1353518}{10.1080/20445911.2017.1353518}

\hypertarget{ref-blais_trial-by-trial_2015}{}
Blais, C., Harris, M. B., Sinanian, M. H., \& Bunge, S. A. (2015).
Trial-by-trial adjustments in control triggered by incidentally encoded
semantic cues. \emph{The Quarterly Journal of Experimental Psychology},
\emph{68}, 1920--1930.
doi:\href{https://doi.org/10.1080/17470218.2014.1000346}{10.1080/17470218.2014.1000346}

\hypertarget{ref-brady_visual_2013}{}
Brady, T. F., Konkle, T., Gill, J., Oliva, A., \& Alvarez, G. A. (2013).
Visual long-term memory has the same limit on fidelity as visual working
memory. \emph{Psychological Science}, \emph{24}, 981--990.
doi:\href{https://doi.org/10.1177/0956797612465439}{10.1177/0956797612465439}

\hypertarget{ref-brewer_dual_1988}{}
Brewer, M. B. (1988). A dual process model of impression formation. In
R. S. Wyer Jr \& T. K. Srull (Eds.), \emph{A dual-process model of
impression formation: Advances in social cognition} (Vol. 1, pp. 1--36).
Hillsdale, NJ: Erlbaum.

\hypertarget{ref-brosowsky_context-specific_2016}{}
Brosowsky, N. P., \& Crump, M. J. (2016). Context-specific attentional
sampling: Intentional control as a pre-requisite for contextual control.
\emph{Consciousness and Cognition}, \emph{44}, 146--160.
doi:\href{https://doi.org/10.1016/j.concog.2016.07.001}{10.1016/j.concog.2016.07.001}

\hypertarget{ref-brosowsky_memory-guided_2018}{}
Brosowsky, N. P., \& Crump, M. J. C. (2018). Memory-guided selective
attention: Single experiences with conflict have long-lasting effects on
cognitive control. \emph{Journal of Experimental Psychology: General}.
doi:\href{https://doi.org/10.1037/xge0000431}{10.1037/xge0000431}

\hypertarget{ref-bugg_conflict-triggered_2014}{}
Bugg, J. (2014). Conflict-triggered top-down control: Default mode, last
resort, or no such thing? \emph{Journal of Experimental Psychology:
Learning, Memory, and Cognition}, \emph{40}(2), 567.

\hypertarget{ref-bugg_support_2012}{}
Bugg, J., \& Crump, M. J. (2012). In support of a distinction between
voluntary and stimulus-driven control: A review of the literature on
proportion congruent effects. \emph{Frontiers in Psychology}, \emph{3},
367.
doi:\href{https://doi.org/10.3389/fpsyg.2012.00367}{10.3389/fpsyg.2012.00367}

\hypertarget{ref-bugg_converging_2013}{}
Bugg, J., \& Hutchison, K. A. (2013). Converging evidence for control of
color--word Stroop interference at the item level. \emph{Journal of
Experimental Psychology: Human Perception and Performance},
\emph{39}(2), 433.

\hypertarget{ref-bugg_multiple_2008}{}
Bugg, J., Jacoby, L. L., \& Toth, J. P. (2008). Multiple levels of
control in the Stroop task. \emph{Memory \& Cognition}, \emph{36},
1484--1494.
doi:\href{https://doi.org/10.3758/MC.36.8.1484}{10.3758/MC.36.8.1484}

\hypertarget{ref-canadas_social_2013}{}
Cañadas, E., Rodríguez-Bailón, R., Milliken, B., \& Lupiáñez, J. (2013).
Social categories as a context for the allocation of attentional
control. \emph{Journal of Experimental Psychology: General}, \emph{142},
934--943. doi:\href{https://doi.org/10.1037/a0029794}{10.1037/a0029794}

\hypertarget{ref-corballis_independent_2003}{}
Corballis, P. M., \& Gratton, G. (2003). Independent control of
processing strategies for different locations in the visual field.
\emph{Biological Psychology}, \emph{64}, 191--209.
doi:\href{https://doi.org/10.1016/S0301-0511(03)00109-1}{10.1016/S0301-0511(03)00109-1}

\hypertarget{ref-crisp_multiple_2007}{}
Crisp, R. J., \& Hewstone, M. (2007). Multiple social categorization.
\emph{Advances in Experimental Social Psychology}, \emph{39}, 163--254.

\hypertarget{ref-crump_learning_2016}{}
Crump, M. J. (2016). Learning to selectively attend from
context-specific attentional histories: A demonstration and some
constraints. \emph{Canadian Journal of Experimental Psychology/Revue
Canadienne de Psychologie Expérimentale}, \emph{70}, 59--77.
doi:\href{https://doi.org/10.1037/cep0000066}{10.1037/cep0000066}

\hypertarget{ref-crump_flexibility_2009}{}
Crump, M. J. C., \& Milliken, B. (2009). The flexibility of
context-specific control: Evidence for context-driven generalization of
item-specific control settings. \emph{The Quarterly Journal of
Experimental Psychology}, \emph{62}, 1523--1532.
doi:\href{https://doi.org/10.1080/17470210902752096}{10.1080/17470210902752096}

\hypertarget{ref-crump_context-dependent_2018}{}
Crump, M. J. C., Milliken, B., Leboe-McGowan, J., Leboe-McGowan, L., \&
Gao, X. (2018). Context-Dependent Control of Attention Capture: Evidence
From Proportion Congruent Effects. \emph{Canadian Journal of
Experimental Psychology= Revue Canadienne de Psychologie Experimentale}.

\hypertarget{ref-crump_reproducing_2017}{}
Crump, M. J., Brosowsky, N. P., \& Milliken, B. (2017). Reproducing the
location-based context-specific proportion congruent effect for
frequency unbiased items: A reply to Hutcheon and Spieler (2016).
\emph{The Quarterly Journal of Experimental Psychology}, \emph{70},
1792--1807.
doi:\href{https://doi.org/10.1080/17470218.2016.1206130}{10.1080/17470218.2016.1206130}

\hypertarget{ref-crump_context-specific_2006}{}
Crump, M. J., Gong, Z., \& Milliken, B. (2006). The context-specific
proportion congruent Stroop effect: Location as a contextual cue.
\emph{Psychonomic Bulletin \& Review}, \emph{13}, 316--321.
doi:\href{https://doi.org/10.3758/BF03193850}{10.3758/BF03193850}

\hypertarget{ref-crump_context-specific_2008}{}
Crump, M. J., Vaquero, J. M., \& Milliken, B. (2008). Context-specific
learning and control: The roles of awareness, task relevance, and
relative salience. \emph{Consciousness and Cognition}, \emph{17},
22--36.
doi:\href{https://doi.org/10.1016/j.concog.2007.01.004}{10.1016/j.concog.2007.01.004}

\hypertarget{ref-devine_stereotypes_1989}{}
Devine, P. G. (1989). Stereotypes and prejudice: Their automatic and
controlled components. \emph{Journal of Personality and Social
Psychology}, \emph{56}(1), 5.

\hypertarget{ref-egner_creatures_2014}{}
Egner, T. (2014). Creatures of habit (and control): A multi-level
learning perspective on the modulation of congruency effects.
\emph{Frontiers in Psychology}, \emph{5}, 1247.
doi:\href{https://doi.org/10.3389/fpsyg.2014.01247}{10.3389/fpsyg.2014.01247}

\hypertarget{ref-eriksen_effects_1974}{}
Eriksen, B. A., \& Eriksen, C. W. (1974). Effects of noise letters upon
the identification of a target letter in a nonsearch task.
\emph{Perception \& Psychophysics}, \emph{16}, 143--149.
doi:\href{https://doi.org/10.3758/BF03203267}{10.3758/BF03203267}

\hypertarget{ref-fischer_predicting_2015}{}
Fischer, R., \& Dreisbach, G. (2015). Predicting high levels of
multitasking reduces between-tasks interactions. Retrieved from
\url{http://psycnet.apa.org/psycinfo/2015-47211-001/}

\hypertarget{ref-fischer_context-sensitive_2014}{}
Fischer, R., Gottschalk, C., \& Dreisbach, G. (2014). Context-sensitive
adjustment of cognitive control in dual-task performance. \emph{Journal
of Experimental Psychology: Learning, Memory, and Cognition},
\emph{40}(2), 399.

\hypertarget{ref-fiske_continuum_1990}{}
Fiske, S. T., \& Neuberg, S. L. (1990). A continuum of impression
formation, from category-based to individuating processes: Influences of
information and motivation on attention and interpretation. In
\emph{Advances in experimental social psychology} (Vol. 23, pp. 1--74).
Elsevier.

\hypertarget{ref-hutcheon_limits_2017}{}
Hutcheon, T. G., \& Spieler, D. H. (2017). Limits on the
generalizability of context-driven control. \emph{Quarterly Journal of
Experimental Psychology}, \emph{70}(7), 1292--1304.
doi:\href{https://doi.org/10.1080/17470218.2016.1182193}{10.1080/17470218.2016.1182193}

\hypertarget{ref-hubner_location-specific_2016}{}
Hübner, R., \& Mishra, S. (2016). Location-specific attentional control
is also possible in the Simon task. \emph{Psychonomic Bulletin \&
Review}, \emph{23}, 1867--1872.
doi:\href{https://doi.org/10.3758/s13423-016-1057-y}{10.3758/s13423-016-1057-y}

\hypertarget{ref-king_priming_2012}{}
King, J. A., Korb, F. M., \& Egner, T. (2012). Priming of control:
Implicit contextual cuing of top-down attentional set. \emph{The Journal
of Neuroscience}, \emph{32}, 8192--8200.
doi:\href{https://doi.org/10.1523/JNEUROSCI.0934-12.2012}{10.1523/JNEUROSCI.0934-12.2012}

\hypertarget{ref-logan_attention_1980}{}
Logan, G. D. (1980). Attention and automaticity in Stroop and priming
tasks: Theory and data. \emph{Cognitive Psychology}, \emph{12}(4),
523--553.

\hypertarget{ref-logan_what_1998}{}
Logan, G. D. (1998). What is learned during automatization? II.
obligatory encoding of spatial location. \emph{Journal of Experimental
Psychology: Human Perception and Performance}, \emph{24}(6), 1720.

\hypertarget{ref-logan_when_1979}{}
Logan, G. D., \& Zbrodoff, N. J. (1979). When it helps to be misled:
Facilitative effects of increasing the frequency of conflicting stimuli
in a Stroop-like task. \emph{Memory \& Cognition}, \emph{7}, 166--174.
doi:\href{https://doi.org/10.3758/BF03197535}{10.3758/BF03197535}

\hypertarget{ref-logan_strategies_1984}{}
Logan, G. D., Zbrodoff, N. J., \& Williamson, J. (1984). Strategies in
the color-word Stroop task. \emph{Bulletin of the Psychonomic Society},
\emph{22}(2), 135--138.

\hypertarget{ref-lowe_selective_1982}{}
Lowe, D. G., \& Mitterer, J. O. (1982). Selective and divided Attention
in a Stroop task. \emph{Canadian Journal of Psychology/Revue Canadienne
de Psychologie}, \emph{36}(4), 684.

\hypertarget{ref-ma_chicago_2015}{}
Ma, D. S., Correll, J., \& Wittenbrink, B. (2015). The Chicago face
database: A free stimulus set of faces and norming data. \emph{Behavior
Research Methods}, \emph{47}(4), 1122--1135.
doi:\href{https://doi.org/10.3758/s13428-014-0532-5}{10.3758/s13428-014-0532-5}

\hypertarget{ref-macrae_social_2001}{}
Macrae, C. N., \& Bodenhausen, G. V. (2001). Social cognition:
Categorical person perception. \emph{British Journal of Psychology},
\emph{92}(1), 239--255.

\hypertarget{ref-mayr_spatial_1996}{}
Mayr, U. (1996). Spatial attention and implicit sequence learning:
Evidence for independent learning of spatial and nonspatial sequences.
\emph{Journal of Experimental Psychology: Learning, Memory, and
Cognition}, \emph{22}(2), 350.

\hypertarget{ref-mayr_outsourcing_2007}{}
Mayr, U., \& Bryck, R. L. (2007). Outsourcing control to the
environment: Effects of stimulus/response locations on task selection.
\emph{Psychological Research}, \emph{71}(1), 107--116.
doi:\href{https://doi.org/10.1007/s00426-005-0039-x}{10.1007/s00426-005-0039-x}

\hypertarget{ref-schmidt_stroop_2008}{}
Schmidt, J. R., \& Besner, D. (2008). The Stroop effect: Why proportion
congruent has nothing to do with congruency and everything to do with
contingency. \emph{Journal of Experimental Psychology: Learning, Memory,
and Cognition}, \emph{34}, 514--523.
doi:\href{https://doi.org/10.1037/0278-7393.34.3.514}{10.1037/0278-7393.34.3.514}

\hypertarget{ref-stroop_studies_1935}{}
Stroop, J. R. (1935). Studies of interference in serial verbal
reactions. \emph{Journal of Experimental Psychology}, \emph{18}, 643.
doi:\href{https://doi.org/10.1037/h0054651}{10.1037/h0054651}

\hypertarget{ref-tottenham_nimstim_2009}{}
Tottenham, N., Tanaka, J. W., Leon, A. C., McCarry, T., Nurse, M., Hare,
T. A., \ldots{} Nelson, C. (2009). The NimStim set of facial
expressions: Judgments from untrained research participants.
\emph{Psychiatry Research}, \emph{168}(3), 242--249.
doi:\href{https://doi.org/10.1016/j.psychres.2008.05.006}{10.1016/j.psychres.2008.05.006}

\hypertarget{ref-van_selst_solution_1994}{}
Van Selst, M., \& Jolicoeur, P. (1994). A solution to the effect of
sample size on outlier elimination. \emph{The Quarterly Journal of
Experimental Psychology Section A}, \emph{47}, 631--650.
doi:\href{https://doi.org/10.1080/14640749408401131}{10.1080/14640749408401131}

\hypertarget{ref-vietze_context_2009}{}
Vietze, I., \& Wendt, M. (2009). Context specificity of conflict
frequency-dependent control. \emph{The Quarterly Journal of Experimental
Psychology}, \emph{62}, 1391--1400.
doi:\href{https://doi.org/10.1080/17470210802426908}{10.1080/17470210802426908}

\hypertarget{ref-weidler_transfer_2016}{}
Weidler, B. J., \& Bugg, J. (2016). Transfer of location-specific
control to untrained locations. \emph{The Quarterly Journal of
Experimental Psychology}, \emph{69}, 2202--2217.
doi:\href{https://doi.org/10.1080/17470218.2015.1111396}{10.1080/17470218.2015.1111396}

\hypertarget{ref-weidler_attentional_2018}{}
Weidler, B. J., Dey, A., \& Bugg, J. M. (2018). Attentional control
transfers beyond the reference frame. \emph{Psychological Research},
1--14.

\hypertarget{ref-west_effects_1998}{}
West, R., \& Baylis, G. C. (1998). Effects of increased response
dominance and contextual disintegration on the Stroop interference
effect in older adults. \emph{Psychology and Aging}, \emph{13}(2), 206.

\endgroup

\end{document}