### 2018-12-09 redoc_manuscript.Rmd

---
title             : "Contextual recruitment of selective attention can be updated via changes in task-relevance"
shorttitle        : "Contextual recruitment of selective attention"

author: 
  - name          : "Nicholaus P. Brosowsky"
    affiliation   : "1"
    corresponding : yes    #  Define only one corresponding author
    address       : "The Graduate Center, CUNY, 365 5th Ave, New York, NY 10016"
    email         : "nbrosowsky@gradcenter.cuny.edu"
  - name          : "Matthew J. C. Crump"
    affiliation   : "2"

affiliation:
  - id            : "1"
    institution   : "The Graduate Center of the City University of New York"
  - id            : "2"
    institution   : "Brooklyn College of the City University of New York"

abstract: |
  Evidence across a wide variety of attention paradigms shows that environmental cues can trigger adjustments to ongoing priorities for attending to relevant and irrelevant information. This context-specific control over attention suggests that cognitive control can be both automatic and flexible. For instance, in selective attention tasks, congruency effects are larger for items that appear in a context associated with infrequent conflict than in a context associated with frequent conflict. Since the to-be-presented context cannot be predicted or prepared for in advance, attention is assumed to be rapidly updated on-the-fly, triggered by the currently presented context. Context-specific control exemplifies how learning and memory processes can influence attention to enable cognitive flexibility. However, what determines the use of previously learned associations still remains unclear. In the current study, we examined whether task-relevance would influence the learning and use of context cues in a flanker task. Using a secondary counting task, context dimensions associated with differing levels of conflict were made task-relevant or -irrelevant across the experiment. In short, we found that making new contextual information task-relevant caused participants to ignore a previously learned context-attention association and adopt a new context-specific control strategy; all without changing the experimental stimuli. This result suggests that task-relevance is a key determinant of context-specific control.

keywords          : "Attention, Memory, Cognitive Control, Conflict Adaptation, Context-specific, Task Relevance"
wordcount         : "4402"

bibliography      : ["task-relevance-cspc.bib"]

figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : no
mask              : no

#class             : doc
lang              : english

class             : "man"
# output:
#  bookdown::html_document2:
#     toc: true
#     toc_float: true
#     collapsed: false
#     number_sections: false
#     toc_depth: 2
#     code_folding: hide
#     css: webpaper.css

#output: 
#  word_document

output:
#  papaja::apa6_word: default
#  papaja::apa6_pdf: default
   redoc::rdocx_reversible

knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_format = "all") })

header-includes:
  - \usepackage{setspace}\singlespacing 
  - \AtBeginEnvironment{tabular}{\onehalfspacing}
  - \AtBeginEnvironment{lltable}{\onehalfspacing}
  - \captionsetup[table]{font={stretch=1.5}}
  - \captionsetup[figure]{font={stretch=1}}
  - \setlength{\skip\footins}{1cm}
  - \setlength{\footnotesep}{0.5cm}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage[table]{xcolor}
  - \usepackage{wrapfig}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{makecell}
  - \usepackage{setspace}\singlespacing 
  - \AtBeginEnvironment{tabular}{\onehalfspacing}
  - \AtBeginEnvironment{lltable}{\onehalfspacing}
  - \newenvironment{zeroindent}
     {\par\setlength{\parindent}{0pt}}
     {\par}

---

```{r setup, include=FALSE, eval=TRUE,echo=FALSE}
library("dplyr")
library("afex")
library("papaja")
library("magrittr")
library("kableExtra")
#As of 2018/12/07 apa_print() requires development version of Papaja
#devtools::install_github("crsh/papaja")
library(papaja)
```

Selective attention is commonly investigated using interference paradigms such
as the Stroop (1935) and flanker (Eriksen and Eriksen 1974) tasks, where
participants identify a target while ignoring a response-congruent or
-incongruent distractor. {++ Typically, p ++}{-- P --}erformance is
{-- typically --} better {++ (quicker reaction times and higher accuracy) ++} on
congruent versus incongruent trials and the difference-the congruency
effect-taken as an index of attentional priorities. Large congruency effects are
thought to reflect ineffective filtering of the distracting stimuli whereas
small congruency effects are thought to reflect effective filtering. By probing
factors that systematically alter congruency effects, we can then make
inferences about processes that control attentional filtering. For example,
manipulating the frequency of conflict via the proportion of congruent versus
incongruent trials has shown to influence the size of the congruency effect.
Typically, a high proportion congruent experiment produces large congruency
effects, whereas a low proportion congruent experiment produces small congruency
effects (Logan and Zbrodoff 1979; Lowe and Mitterer 1982; West and Baylis 1998).
This result is usually explained as strategic control, where participants
increase attentional control under high-conflict demands and relax attentional
control under low-conflict demands (Logan 1980; Logan, Zbrodoff, and Williamson
1984; Logan and Zbrodoff 1979; Lowe and Mitterer 1982). Recent work however, has
demonstrated that attentional control is not only adjusted by top-down
regulation, but can also be triggered automatically by environmental cues
(Brosowsky and Crump 2018; Bugg and Crump 2012; Egner 2014; Fischer and
Dreisbach 2015; King, Korb, and Egner 2012; Mayr and Bryck 2007).

For example, Crump, Gong, and Milliken (Crump, Gong, and Milliken 2006; see
also, Corballis and Gratton 2003) presented Stroop stimuli in one of two
randomly chosen locations and manipulated the frequency of conflict associated
with each location. One location was associated with a high frequency of
conflict (25% congruent trials) and the other with a low frequency of conflict
(75% congruent trials). Overall, the proportion of congruent trials was 50% and
randomized such that the upcoming location could not be predicted. Even so,
congruency effects were shown to be smaller for trials where the stimulus
appeared in the high conflict location as compared to the low conflict location.
This effect, known now as the context-specific proportion congruent effect
(CSPC), has now been replicated in a number of different selective attention
paradigms (e.g., Alards-Tomalin, Brosowsky, and Mondor 2017; Blais et al. 2015;
Bugg 2014; Crump et al. 2018; Crump 2016; Fischer, Gottschalk, and Dreisbach
2014; Hübner and Mishra 2016).

Critical evidence however, that CSPC effects reflect context-specific control
rather than other non-control learning processes (e.g., Schmidt and Besner
2008), comes from work showing that CSPC effects can transfer to frequency
unbiased items (Brosowsky and Crump 2016; Crump and Milliken 2009; Weidler, Dey,
and Bugg 2018; Weidler and Bugg 2016; though, see Hutcheon and Spieler 2017).
Crump and Milliken (2009), for example, divided Stroop items into two mutually
exclusive sets (e.g., RED/GREEN and BLUE/YELLOW). One set was defined as the
frequency biased set, and presented with 75% congruency in one location, and 25%
congruency in the other. The second set however, was presented with 50%
congruency in both locations. Nevertheless, they found smaller congruency
effects for unbiased items presented in the high conflict location as compared
to the low conflict location.

One explanation for CSPC effects is that the repeated application of attentional
priorities in a particular context creates an associative link in episodic
memory between the attentional control procedures and the contextual information
(Abrahamse et al. 2016; Brosowsky and Crump 2018; Crump 2016; Egner 2014). Once
the associative link is established, processing the context is assumed to
trigger the retrieval of previous experiences, automatically reinstating the
associated attentional priorities. Although this explanation seems plausible, it
remains largely unclear how such associations are established and perhaps more
importantly, it remains unclear what determines which learned associations will
be used to guide attention at any given moment.

The importance of the latter issue becomes increasingly clear if we speculate
about context-specificity in more complex, real-world environments. Consider
myself, right now, writing this article in a coffee shop. The coffee shop is a
complex environment that contains a multitude of contextual cues (e.g., pictures
on the wall, chairs, coffee, laptop), each of which I have had many different
experiences applying attentional control. Although my laptop might cue
attentional processing helpful for finishing this article, other features of the
coffee shop might cue other, unhelpful attentional processing related to other
experiences like socializing and people-watching. This is all to say that our
environments, and experiences within them, afford many different associative
relationships between cognitive processing and environmental cues. What then
determines which associative relationships will be used to guide attention (or
even if such associative relationships are used at all)?

Task-relevance might be one factor that is important for establishing and using
context-attention associations. This proposal might seem counter-intuitive as it
is typically assumed that the context, within selective attention tasks at
least, is completely irrelevant to the task. In fact, referring to such stimulus
dimensions as the "context" carries with it an implication that the feature is
background information, irrelevant to the task-at-hand, and should be ignored by
the participant. A closer look at CSPC designs however, suggests that the
context cues may not be as irrelevant as implied. For example, the most commonly
used context cue is location (Brosowsky and Crump 2016; Corballis and Gratton
2003; Crump, Gong, and Milliken 2006; Crump, Vaquero, and Milliken 2008; Crump
2016; Weidler, Dey, and Bugg 2018; Weidler and Bugg 2016). Although in some
sense, the location is irrelevant for identifying the color of a word, it seems
disingenuous to consider it completely task-irrelevant. In order to identify the
target, the participant must first identify and orient to its location on the
screen. In this is sense, it is relevant to their ability to complete the task
and furthermore, location information may receive priority during encoding
(Logan 1998; Mayr 1996). Similar arguments could be made about other context
features that are inherent to processing the target stimulus, like font-type
(Bugg, Jacoby, and Toth 2008) and color (Vietze and Wendt 2009).

To date, we are aware of only one study that has looked at task-relevance within
the CSPC literature. Crump, Vaquero, & Milliken (2008) used a Stroop prime-probe
task where the distractor word is presented first followed by a color patch
probe. Critically, they used the shape of the color patch probe, either a square
or a circle, as the contextual dimension; one associated with a high frequency
of conflict, the other with a low frequency. In one experiment participants were
made aware of the shape-proportion contingencies but given no additional
instructions. In the second, participants were given a secondary task to count
the number of trials that contained a square. The first experiment failed to
find evidence for context-specificity, whereas the second did. Although not
testing task-relevance directly, Cañadas et al. (2013)\] found similar effects.
In this study they used images of male and female faces as context cues in a
flanker task. When given instructions to think of the faces as members of the
gender categories they found CSPC effects, but when given instructions to think
of the faces as individuals they did not. These results suggest that making a
context dimension relevant to the on-going task is important for producing CSPC
effects.

The Current Study
=================

Prior work suggests that the task-relevance of contextual information is
important for establishing new associations between context cues and attentional
priorities. Simply put, CSPC effects emerge when context cues are somehow made
relevant to the task, but do not when made irrelevant (Cañadas et al. 2013;
Crump, Gong, and Milliken 2006; Crump, Vaquero, and Milliken 2008). However, it
still remains unclear how task-relevance may or may not inform the use of
previously learned associations. In the current study, we examined whether
manipulating task-relevance would cause participants to ignore a previously
learned association and adopt a new context-specific control strategy.

A flanker task was used to measure attentional control in an adapted CSPC design
(Crump and Milliken 2009; Crump, Brosowsky, and Milliken 2017). Similar to
previous studies, flanker stimuli were presented in contexts associated with
differing levels of conflict (0%, 50%, or 100% congruent). However, unlike
previous CSPC designs where contexts were defined by a single discriminating
feature (e.g., upper vs. lower screen locations), contexts were defined by two
feature dimensions (see Figure 1A). In one condition, for example, contexts were
defined by object identity (hat or chair) and color (blue or green). Only three
out of the four possible feature combinations were presented, each associated
with a different proportion of congruent trials (0%, 50%, or 100% congruent).
Critically, the frequency unbiased context (50% congruent) always shared a
feature with each of the frequency-biased contexts (0% and 100% congruent). The
frequency biased contexts (0% and 100% congruent) however, did not share any
features (see Figure 1C for an example).

A secondary counting task was used to manipulate the task-relevance of the
context dimensions (Crump and Milliken 2009) which critically, switched halfway
through the experiment. Participants were instructed to keep a running count of
one of the overlapping features (e.g., "count whenever a hat is presented")
which was associated with either a high or low frequency of conflict. Halfway
through the experiment they received new instructions to count the other
overlapping feature (e.g., "count whenever a green item is presented").
Critically, the set of context images remained the same throughout the
experiment (see Figure 1C for an example) and the task-relevance of the context
dimensions was the only aspect of the experiment that changed from the first to
second phases.

The critical measure of interest is the congruency effect produced in the
frequency unbiased context. If, on the one hand, task-relevance has no impact on
the use of context-attention associations we would expect no differences between
the congruency effects when the high-conflict context dimension is made
task-relevant or the low-conflict context dimension is made task-relevant. This
could occur because all three contexts are treated as individual contexts
throughout the whole experiment (e.g., Cañadas et al. 2013). Or it could occur
because associations formed in the first block interfere with learning new
associations in the second (e.g., Brosowsky and Crump 2016). On other hand,
task-relevance might not only dictate how associations are formed, but also
whether learned associations are used. In this case, we would expect smaller
congruency effects when the high-conflict context dimension is made
task-relevant as compared to when the low-conflict context dimension is made
task-relevant. This would demonstrate that participants were able to ignore a
previously learned association and adopt a new context-specific control
strategy.

A secondary goal was to conceptually replicate previous findings and test the
generalizability of any task-relevance effects. Therefore, we included three
conditions that differed only in the kinds of stimuli used to create context
cues. In one condition we used face images where context dimensions were defined
by social categories, similar to Cañadas et al. (2013). However, they found that
social categorization occurred spontaneously and it was unclear what impact that
would have on the current results. Therefore, we included another condition that
used images of simple objects where context dimensions were defined by object
features. Finally, Cañadas et al. (2013) also found that context effects
generalized to novel face images. In our third condition we used non-repeating
face exemplars to examine whether any task-relevance effects would also
generalize in our design or if learning would be image-specific.

Methods
=======

(ref:figure1) Illustration of the stimuli and trial construction. Figure 1A
shows the feature dimensions for each of the conditions. Figure 1B shows an
example trial stimulus containing both the context and flanker images. Figure 1C
shows an example of how feature dimensions could have been assigned to each
proportion congruency (PC) condition, and an example of the secondary task
assignments. The secondary task order, as well as the feature and congruency
assignments were all randomized for each participant.

```{r figure1, echo=FALSE, fig.width=6.7,fig.height=4.51, fig.cap='(ref:figure1)'}
knitr::include_graphics("images/Task.jpg")
```

Participants
------------

All participants were recruited from Amazon Mechanical Turk (AMT) and
compensated \$2.00 for participating. The amount compensated was calculated by
estimating the maximum amount of time required to complete each experiment and
multiplying by \$6.00 per hour. For each experiment the number of HITs (Human
intelligence tasks, an Amazon term for a work-unit) refers to the number of
participants who initiated the study. Participants were included in the study if
they completed all trials. We posted 150 HITs and 144 participants completed all
trials.

Apparatus & Stimuli
-------------------

The experiments were programmed using JavaScript, CSS and HTML. The program
allowed participants to complete task only if they were running Safari, Google
Chrome, or Firefox web browsers. Flanker stimuli consisted of images of five
arrows pointed left or right presented at 250 x 50 pixels (each arrow was 50 x
50 pixels). Context stimuli were constructed using images selected from Brady et
al. (2013) color-rotated to blue and green, and face images from the Chicago
Face Database (Ma, Correll, and Wittenbrink 2015), supplemented with the NimStim
Set of Facial Expressions (Tottenham et al. 2009). The object images were
displayed at 250 x 250 pixels, while the face images were displayed at 250 x 313
pixels. The experiment ran as a pop-up window that filled the entire screen. The
background was white, and stimuli were presented in the center of the screen.

Design. Experiment 1 used a 2x2x3 mixed design with task-relevant context (0%
and 100%) and unbiased-item congruency (congruent and incongruent) as
within-subject factors, and context-type (object, social, and
social/non-repeating) as the between-subjects factor. All three conditions were
constructed using the same general method. The experiment was divided into two
phases. Each phase consisted of 144 flanker trials (48 trials per context), and
13 count response trials for a total of 314 trials. On the count response
trials, participants indicated how many trials they had counted until that
point. The count response trials occurred once for every 12 flanker trials and
was randomly inserted between trial 6 and 12 of each 12-trial block. Each phase
ended with one additional count response trial.

On every flanker trial, participants were presented with flanker stimuli paired
with one of three contexts. Each context was associated with a different
proportion congruency such that two cues were associated with a biased frequency
(0% and 100% proportion congruency), while one was associated with an unbiased
frequency (50% proportion congruency). The feature dimensions and corresponding
context images assigned to each of the biased and unbiased item sets were
randomly determined for each participant. However, context images used for the
frequency biased trials never shared features, while the frequency unbiased
context image always shared a feature with each of the frequency biased context
images (see Figure 1). Additionally, the feature assignments remained the same
throughout phases 1 and 2, and critically, the only change to the task was which
feature the participant was instructed to count (see Figure 1C for an example).

All critical aspects of the task were randomized between participants. This
includes the three chosen context images, the features assigned to proportion
levels, the features assigned to each counting condition, the secondary task
order, and the order of trials.

Procedure
---------

All participants were AMT workers who found the experiment using the AMT system.
The participant recruitment procedure and tasks were approved by the Brooklyn
College Institutional Review Board. Each participant read a short description of
the task and gave consent by pressing a button acknowledging they had read the
displayed consent form. Participants then completed a short demographic survey,
and proceeded to the main task, which was displayed as a pop-up window.
Participants were instructed to identify the direction of the center arrow on
each trial as quickly and accurately as possible by pressing 'z' if the arrow
pointed left, and 'm' if the arrow pointed right. Additionally, they were
instructed to silently keep count of the number of trials that contained a
feature. In the object context condition, they were asked to count trials that
contained a certain color (blue or green) or object-identity (hat or chair) and
in the social context conditions they were asked to count the number of trials
that contained appeared a certain gender (male or female) or race (black or
white). Periodically throughout the experiment, participants were asked to
report how many trials they had counted until that point and to restart their
count from 0. Halfway through the experiment participants received new
instructions about which feature to count (see Figure 1C). Each trial began with
a blank inter-stimulus interval (ISI) of 400 ms, followed by a fixation cross
presented in the center of the screen for 200 ms, then a second blank ISI of 400
ms. Next, the flanker and context stimuli appeared in the center of screen (the
flanker above the context image; see Figure 1B) and remained on screen until a
response was made. Following a response, accuracy feedback was presented for
1000 ms. A response automatically triggered the next trial. Halfway through the
experiment (157 trials), participants received new instructions about which
feature to count and to press the button on-screen when they were ready to
continue.

Results
=======

```{r analysis}
source("data-analysis/analysis.R")
```

(ref:figure2) Results from experiment 1 showing congruency effects (incongruent
- congruent) for frequency unbiased items in reaction times (left) and error
rates (right) as a function of the task-relevant context (100% proportion
congruent versus 0% proportion congruent).

```{r figure2, echo=FALSE,fig.width=6.875, fig.height= 4, fig.cap='(ref:figure2)'}
figure2
```

Participants with mean error rates greater than 25% were excluded from the
analyses, eliminating `r N_subjects_removed` participants. For all remaining participants,
the correct RTs from frequency unbiased trials in each condition were submitted
to an outlier removal procedure. The non-recursive Van Selst and Jolicoeur
outlier removal procedure was applied after removing response times greater than
3000 ms (Van Selst and Jolicoeur 1994). This procedure removed `r round(percent_removed*100,digits = 2)`% of
the total observations.

The primary question of interest was whether the task-relevance of context
features associated with different levels of conflict would influence the size
of the congruency effect for frequency unbiased items (see Table 1). To that
end, mean correct RTs and mean error rates from frequency unbiased trials were
submitted to a mixed analysis of variance (ANOVA) with task-relevant context (0%
and 100% PC) and unbiased-item congruency (congruent and incongruent) as
within-subject factors, and context-type (object, social, and
social/non-repeating) as the between-subjects factor.

The results of the RT analysis revealed the critical, two-way interaction
between task-relevant context and unbiased-item congruency to be significant,
$F(1,128) = 8.30$`r RT_ANOVA$Task_Relevant_Context_Congruency` $\text{MSE} = 1,608.54$`r RT_ANOVA$Task_Relevant_Context_Congruency`
$p = .005$`r RT_ANOVA$Task_Relevant_Context_Congruency` ${\widehat{\eta}}_{p}^{2} = .061$, demonstrating smaller
congruency effects when the context dimension associated with high conflict was
made task-relevant. Additionally, the three-way interaction between the
task-relevant context, unbiased-item congruency, and context-type was
non-significant, $F(2,128) = 0.00$`r RT_ANOVA$Condition_Task_Relevant_Context_Congruency`
$\text{MSE} = 1,608.54$`r RT_ANOVA$Condition_Task_Relevant_Context_Congruency` $p = .996$`r RT_ANOVA$Condition_Task_Relevant_Context_Congruency`
${\widehat{\eta}}_{p}^{2} = .000$, showing no evidence for differences between
the CSPC effects across context-type manipulations.

Similarly, the corresponding error analysis also resulted in a significant
two-way interaction between the task-relevant context and unbiased-item
congruency, $F(1,128) = 6.72$`r ACC_ANOVA$Task_Relevant_Context_Congruency` $\text{MSE} = 13.93$`r ACC_ANOVA$Task_Relevant_Context_Congruency`
$p = .011$`r ACC_ANOVA$Task_Relevant_Context_Congruency` ${\widehat{\eta}}_{p}^{2} = .050$, and a non-significant
three-way interaction between task-relevant context, unbiased-item congruency,
and context-type, $F(2,128) = 0.92$`r ACC_ANOVA$Condition_Task_Relevant_Context_Congruency` $\text{MSE} = 13.93$`r ACC_ANOVA$Condition_Task_Relevant_Context_Congruency`
$p = .400$`r ACC_ANOVA$Condition_Task_Relevant_Context_Congruency` ${\widehat{\eta}}_{p}^{2} = .014$. Therefore, the
congruency effects in error rates were also smaller when the context dimension
associated with high conflict was made task-relevant, corroborating the results
of the reaction time analysis.

General Discussion
==================

The aim of the current study was to investigate whether task-relevance plays a
role in the contextual recruitment of selective attention. Specifically, we
tested whether manipulating the relative task-relevance of context cues could
cause participants to ignore a previously learned context-association and apply
a new association. We created a frequency unbiased context cue that shared
features with two frequency biased contexts and included a feature-counting task
to manipulate the task-relevance of context dimensions across two blocks of
trials. Critically, halfway through the experiment participants received new
instructions changing the task-relevant feature from one frequency biased cue to
the other.

The key finding was that the congruency effects for the frequency unbiased items
were significantly larger when the low-conflict context was made task-relevant
as compared to when the high-conflict context was made task-relevant. This
result is consistent with prior CSPC effects and, like the previous work,
suggests that the context cues triggered rapid adjustments to attentional
control (Crump and Milliken 2009). However, unlike prior studies, we were able
to experimentally manipulate the CSPC effect across blocks of trials without
changing any of the physical properties of the stimuli. This novel finding
demonstrates that participants were able to learn and apply one
context-attention association in the first phase, and subsequently ignore that
association to learn a new association in the second phase.

This result implicates an important role for task-relevance in producing CSPC
phenomena. Crump, Vaquero, and Milliken (2008) used shapes as context cues in a
prime-probe Stroop task and did not find CSPC effects until the context cues
were made task-relevant. Similarly, Cañadas et al. (2013) eliminated the CSPC
effect by making the contextual cue effectively unrelated to the task. These
studies suggest that task-relevance plays an important role in establishing
associations between contextual information and attentional priorities to
produce CSPC effects. Our finding extends this work in two important ways.
First, we show that changing the task-relevance of the presented cues
corresponded with a change in attentional control in the predicted direction.
This demonstrates that task-relevance is also a key determinant of whether a
previously learned attention-context association will be used or ignored in
favor of a new association. Second, we show that task-relevance allowed
participants resolve competition between two competing contextual cues,
responding on the basis of one at the expense of the other. To our knowledge,
this is the first demonstration that CSPC effects can be produced when there are
multiple, overlapping contextual cues available.

In light of prior work, we take this as evidence that the contextual recruitment
of selective attention, although likely implicit, is not obligatory (e.g.,
Brosowsky and Crump 2016), requiring that environmental information be
incorporated into the task representation. Similarly, there appears to be
flexibility in which environmental features are selected and used to guide
attention, which can be rapidly updated depending on the task-relevance of those
features. Such a result lends some insight into how context-specific control
might operate in more complex, real-world environments, where there is an
over-abundance of environmental features that afford many different learned
associations. From a theoretical perspective, this result is consistent
memory-based accounts of CSPC phenomena (Brosowsky and Crump 2018; Bugg and
Hutchison 2013; Crump et al. 2018; Crump and Milliken 2009; Crump, Brosowsky,
and Milliken 2017). Under this view, a memory process encodes attentional
priorities in the representation of individual experiences and, as a result,
become associated with the environment where they were used. The subsequent
reoccurrence of a prior context triggers the retrieval and reinstatement of
those attentional priorities. Our results show however, that all the features of
the environment may not be treated equally and that only task-relevant features
are used to probe memory and guide attention.

Another key result of this study concerned the different stimuli used as context
cues. Across the three conditions, we varied the type of context image and
dimensions. We manipulated the type of image presented, including both objects
(identity and color dimensions) and faces (gender and racial dimensions). We
also manipulated whether a single set of three repeating images were presented
(object and social) or a set of non-repeating images were presented
(social/non-repeating). Across all three conditions, we found no evidence that
using different stimuli had an influence on the size or direction of the CSPC
demonstrating generalizability of this phenomenon. Furthermore, CSPC effects
were present even when using non-repeating images which suggests that
context-dependency did rely on image-specific associations but higher-order,
learned categorical information.

Traditional models of person perception posit that social categories are
automatically activated in the presence of social stimuli (e.g., Brewer 1988;
Devine 1989; Fiske and Neuberg 1990). Cañadas et al. (2013) however, found that
directing participants to think about faces in terms of individual features
eliminated context-specific attention effect and suggested that momentary
motivations may influence the automaticity of social categorization. Our results
add to this literature by observing the influence of momentary motivations
(i.e., task-relevance) when there is competition between two salient social
categories. Specifically, we found that participants could categorize on the
basis of one social cue at the expense of the other, and subsequently switch
between them. These findings may speak to issues of automaticity in social
categorization (Macrae and Bodenhausen 2001) as well as understanding how the
situational context can prime one social identity over another (Crisp and
Hewstone 2007). Furthermore, we found no evidence for differences between the
social and object conditions. This suggests that categorization within the CSPC
task is quickly and easily learned when the task supports such learning and is
non-unique to social stimuli.

In sum, our results provide new evidence that changes in task-relevance can
update the contextual recruitment of selective attention in a CSPC flanker task.
Importantly, we show that making one feature task-relevant produced a CSPC
effect, even in the presence of a competing contextual cue. Moreover, changing
the task-relevant context-cue midway through the experiment was accompanied by
predictable changes in the congruency effects. This effect was found to be
generalizable across two different kinds of stimuli and occurred even when using
non-repeating images. These results implicate an important role for
task-relevance in producing context-dependency and are consistent with
memory-based accounts of CSPC phenomena.

References
==========

Abrahamse, Elger, Senne Braem, Wim Notebaert, and Tom Verguts. 2016. "Grounding
Cognitive Control in Associative Learning." *Psychological Bulletin* 142:
693--728. <https://doi.org/10.1037/bul0000047>.

Alards-Tomalin, Doug, Nicholaus P. Brosowsky, and Todd A. Mondor. 2017.
"Auditory Statistical Learning: Predictive Frequency Information Affects the
Deployment of Contextually Mediated Attentional Resources on Perceptual Tasks."
*Journal of Cognitive Psychology* 0 (0): 1--11.
<https://doi.org/10.1080/20445911.2017.1353518>.

Blais, Chris, Michael B. Harris, Michael H. Sinanian, and Silvia A. Bunge. 2015.
"Trial-by-Trial Adjustments in Control Triggered by Incidentally Encoded
Semantic Cues." *The Quarterly Journal of Experimental Psychology* 68: 1920--30.
<https://doi.org/10.1080/17470218.2014.1000346>.

Brady, Timothy F., Talia Konkle, Jonathan Gill, Aude Oliva, and George A.
Alvarez. 2013. "Visual Long-Term Memory Has the Same Limit on Fidelity as Visual
Working Memory." *Psychological Science* 24: 981--90.
<https://doi.org/10.1177/0956797612465439>.

Brewer, M. B. 1988. "A Dual Process Model of Impression Formation." In *A
Dual-Process Model of Impression Formation: Advances in Social Cognition*,
edited by Robert S. Wyer Jr and Thomas K. Srull, 1:1--36. Hillsdale, NJ:
Erlbaum.

Brosowsky, Nicholaus P., and Matthew J. C. Crump. 2018. "Memory-Guided Selective
Attention: Single Experiences with Conflict Have Long-Lasting Effects on
Cognitive Control." *Journal of Experimental Psychology: General*.
<https://doi.org/10.1037/xge0000431>.

Brosowsky, Nicholaus P., and Matthew J.C. Crump. 2016. "Context-Specific
Attentional Sampling: Intentional Control as a Pre-Requisite for Contextual
Control." *Consciousness and Cognition* 44: 146--60.
<https://doi.org/10.1016/j.concog.2016.07.001>.

Bugg, J.M. 2014. "Conflict-Triggered Top-down Control: Default Mode, Last
Resort, or No Such Thing?" *Journal of Experimental Psychology: Learning,
Memory, and Cognition* 40 (2): 567.

Bugg, J.M., and Matthew J.C. Crump. 2012. "In Support of a Distinction Between
Voluntary and Stimulus-Driven Control: A Review of the Literature on Proportion
Congruent Effects." *Frontiers in Psychology* 3: 367.
<https://doi.org/10.3389/fpsyg.2012.00367>.

Bugg, J.M., and Keith A. Hutchison. 2013. "Converging Evidence for Control of
Color--Word Stroop Interference at the Item Level." *Journal of Experimental
Psychology: Human Perception and Performance* 39 (2): 433.

Bugg, J.M., Larry L. Jacoby, and Jeffrey P. Toth. 2008. "Multiple Levels of
Control in the Stroop Task." *Memory & Cognition* 36: 1484--94.
<https://doi.org/10.3758/MC.36.8.1484>.

Cañadas, Elena, Rosa Rodríguez-Bailón, Bruce Milliken, and Juan Lupiáñez. 2013.
"Social Categories as a Context for the Allocation of Attentional Control."
*Journal of Experimental Psychology: General* 142: 934--43.
<https://doi.org/10.1037/a0029794>.

Corballis, Paul M., and Gabriele Gratton. 2003. "Independent Control of
Processing Strategies for Different Locations in the Visual Field." *Biological
Psychology* 64: 191--209. <https://doi.org/10.1016/S0301-0511(03)00109-1>.

Crisp, Richard J., and Miles Hewstone. 2007. "Multiple Social Categorization."
*Advances in Experimental Social Psychology* 39: 163--254.

Crump, Matthew J.C. 2016. "Learning to Selectively Attend from Context-Specific
Attentional Histories: A Demonstration and Some Constraints." *Canadian Journal
of Experimental Psychology/Revue Canadienne de Psychologie Expérimentale* 70:
59--77. <https://doi.org/10.1037/cep0000066>.

Crump, Matthew J.C., Nicholaus P. Brosowsky, and Bruce Milliken. 2017.
"Reproducing the Location-Based Context-Specific Proportion Congruent Effect for
Frequency Unbiased Items: A Reply to Hutcheon and Spieler (2016)." *The
Quarterly Journal of Experimental Psychology* 70: 1792--1807.
<https://doi.org/10.1080/17470218.2016.1206130>.

Crump, Matthew J.C., Zhiyu Gong, and Bruce Milliken. 2006. "The Context-Specific
Proportion Congruent Stroop Effect: Location as a Contextual Cue." *Psychonomic
Bulletin & Review* 13: 316--21. <https://doi.org/10.3758/BF03193850>.

Crump, Matthew J. C., B. Milliken, J. Leboe-McGowan, L. Leboe-McGowan, and X.
Gao. 2018. "Context-Dependent Control of Attention Capture: Evidence from
Proportion Congruent Effects." *Canadian Journal of Experimental Psychology=
Revue Canadienne de Psychologie Experimentale*.

Crump, Matthew J. C., and Bruce Milliken. 2009. "The Flexibility of
Context-Specific Control: Evidence for Context-Driven Generalization of
Item-Specific Control Settings." *The Quarterly Journal of Experimental
Psychology* 62: 1523--32. <https://doi.org/10.1080/17470210902752096>.

Crump, Matthew J.C., Joaquín M.M. Vaquero, and Bruce Milliken. 2008.
"Context-Specific Learning and Control: The Roles of Awareness, Task Relevance,
and Relative Salience." *Consciousness and Cognition* 17: 22--36.
<https://doi.org/10.1016/j.concog.2007.01.004>.

Devine, Patricia G. 1989. "Stereotypes and Prejudice: Their Automatic and
Controlled Components." *Journal of Personality and Social Psychology* 56 (1):
5.

Egner, Tobias. 2014. "Creatures of Habit (and Control): A Multi-Level Learning
Perspective on the Modulation of Congruency Effects." *Frontiers in Psychology*
5: 1247. <https://doi.org/10.3389/fpsyg.2014.01247>.

Eriksen, Barbara A., and Charles W. Eriksen. 1974. "Effects of Noise Letters
Upon the Identification of a Target Letter in a Nonsearch Task." *Perception &
Psychophysics* 16: 143--49. <https://doi.org/10.3758/BF03203267>.

Fischer, Rico, and Gesine Dreisbach. 2015. "Predicting High Levels of
Multitasking Reduces Between-Tasks Interactions."
<http://psycnet.apa.org/psycinfo/2015-47211-001/>.

Fischer, Rico, Caroline Gottschalk, and Gesine Dreisbach. 2014.
"Context-Sensitive Adjustment of Cognitive Control in Dual-Task Performance."
*Journal of Experimental Psychology: Learning, Memory, and Cognition* 40 (2):
399.

Fiske, Susan T., and Steven L. Neuberg. 1990. "A Continuum of Impression
Formation, from Category-Based to Individuating Processes: Influences of
Information and Motivation on Attention and Interpretation." In *Advances in
Experimental Social Psychology*, 23:1--74. Elsevier.

Hutcheon, Thomas G., and Daniel H. Spieler. 2017. "Limits on the
Generalizability of Context-Driven Control." *Quarterly Journal of Experimental
Psychology* 70 (7): 1292--1304. <https://doi.org/10.1080/17470218.2016.1182193>.

Hübner, Ronald, and Shreyasi Mishra. 2016. "Location-Specific Attentional
Control Is Also Possible in the Simon Task." *Psychonomic Bulletin & Review* 23:
1867--72. <https://doi.org/10.3758/s13423-016-1057-y>.

King, Joseph A., Franziska M. Korb, and Tobias Egner. 2012. "Priming of Control:
Implicit Contextual Cuing of Top-down Attentional Set." *The Journal of
Neuroscience* 32: 8192--8200. <https://doi.org/10.1523/JNEUROSCI.0934-12.2012>.

Logan, Gordon D. 1980. "Attention and Automaticity in Stroop and Priming Tasks:
Theory and Data." *Cognitive Psychology* 12 (4): 523--53.

---------. 1998. "What Is Learned During Automatization? II. Obligatory Encoding
of Spatial Location." *Journal of Experimental Psychology: Human Perception and
Performance* 24 (6): 1720.

Logan, Gordon D., and N. Jane Zbrodoff. 1979. "When It Helps to Be Misled:
Facilitative Effects of Increasing the Frequency of Conflicting Stimuli in a
Stroop-Like Task." *Memory & Cognition* 7: 166--74.
<https://doi.org/10.3758/BF03197535>.

Logan, Gordon D., N. Jane Zbrodoff, and James Williamson. 1984. "Strategies in
the Color-Word Stroop Task." *Bulletin of the Psychonomic Society* 22 (2):
135--38.

Lowe, Douglas G., and John O. Mitterer. 1982. "Selective and Divided Attention
in a Stroop Task." *Canadian Journal of Psychology/Revue Canadienne de
Psychologie* 36 (4): 684.

Ma, Debbie S., Joshua Correll, and Bernd Wittenbrink. 2015. "The Chicago Face
Database: A Free Stimulus Set of Faces and Norming Data." *Behavior Research
Methods* 47 (4): 1122--35. <https://doi.org/10.3758/s13428-014-0532-5>.

Macrae, C. Neil, and Galen V. Bodenhausen. 2001. "Social Cognition: Categorical
Person Perception." *British Journal of Psychology* 92 (1): 239--55.

Mayr, Ulrich. 1996. "Spatial Attention and Implicit Sequence Learning: Evidence
for Independent Learning of Spatial and Nonspatial Sequences." *Journal of
Experimental Psychology: Learning, Memory, and Cognition* 22 (2): 350.

Mayr, Ulrich, and Richard L. Bryck. 2007. "Outsourcing Control to the
Environment: Effects of Stimulus/Response Locations on Task Selection."
*Psychological Research* 71 (1): 107--16.
<https://doi.org/10.1007/s00426-005-0039-x>.

Schmidt, James R., and Derek Besner. 2008. "The Stroop Effect: Why Proportion
Congruent Has Nothing to Do with Congruency and Everything to Do with
Contingency." *Journal of Experimental Psychology: Learning, Memory, and
Cognition* 34: 514--23. <https://doi.org/10.1037/0278-7393.34.3.514>.

Stroop, J. Ridley. 1935. "Studies of Interference in Serial Verbal Reactions."
*Journal of Experimental Psychology* 18: 643.
<https://doi.org/10.1037/h0054651>.

Tottenham, Nim, James W. Tanaka, Andrew C. Leon, Thomas McCarry, Marcella Nurse,
Todd A. Hare, David J. Marcus, Alissa Westerlund, BJ Casey, and Charles Nelson.
2009. "The NimStim Set of Facial Expressions: Judgments from Untrained Research
Participants." *Psychiatry Research* 168 (3): 242--49.
<https://doi.org/10.1016/j.psychres.2008.05.006>.

Van Selst, Mark, and Pierre Jolicoeur. 1994. "A Solution to the Effect of Sample
Size on Outlier Elimination." *The Quarterly Journal of Experimental Psychology
Section A* 47: 631--50. <https://doi.org/10.1080/14640749408401131>.

Vietze, Ina, and Mike Wendt. 2009. "Context Specificity of Conflict
Frequency-Dependent Control." *The Quarterly Journal of Experimental Psychology*
62: 1391--1400. <https://doi.org/10.1080/17470210802426908>.

Weidler, Blaire J., and J.M. Bugg. 2016. "Transfer of Location-Specific Control
to Untrained Locations." *The Quarterly Journal of Experimental Psychology* 69:
2202--17. <https://doi.org/10.1080/17470218.2015.1111396>.

Weidler, Blaire J., Abhishek Dey, and Julie M. Bugg. 2018. "Attentional Control
Transfers Beyond the Reference Frame." *Psychological Research*, 1--14.

West, Robert, and Gordon C. Baylis. 1998. "Effects of Increased Response
Dominance and Contextual Disintegration on the Stroop Interference Effect in
Older Adults." *Psychology and Aging* 13 (2): 206.

Table 1
-------

### Reaction times and error rates from Experiment 1.

\\begin{zeroindent}

```{r tableLatex}

if (knitr::opts_knit$get("rmarkdown.pandoc.to")== "latex"){
source("data-analysis/table1.R")
latexTable1

}
```

\\end{zeroindent}

```{r tableWord, echo=FALSE, out.width = '100%'}
if (knitr::opts_knit$get("rmarkdown.pandoc.to")== "docx"){
  knitr::include_graphics("data-analysis/table1.png")
}

```